---
title: "Data Munging"
output: html_notebook
---

## Libraries
```{r message = FALSE} 
library(tidyverse)
library(skimr)
library(knitr)
library(janitor)
library(glmnet)
library(caret)
library(leaps)
library(rpart)
library(rattle) 

```

## Import Data
```{r message = FALSE}
admit = read_csv("SummerStudentAdmissions2.csv") %>%
  clean_names()

```

## View Data
```{r}
skim(admit) 

```

# Start Initial clean of variables

## Decision Variable
```{r}
# View values of decision
#table(admit$decision)

# remove any odd responses
admit$decision[admit$decision != "Admit" & admit$decision != "Decline" & admit$decision != "Waitlist"] = NA

# Make decision variable a factor
admit$decision = as.factor(admit$decision)

```

## State Variable
```{r}
# View values of state
#table(admit$state)

# Some states need to be capitalized
admit$state[admit$state == "virginia"] = "Virginia"
admit$state[admit$state == "mississippi"] = "Mississippi"

# Make state data a factor
admit$state = as.factor(admit$state)
```

## GPA Variable
```{r}
summary(admit$gpa)

admit %>%
  ggplot(aes(x = gpa)) +
  geom_histogram(fill = "darkblue", alpha = .65, bins = 20) +
  labs(
    title = "Admissions GPAs",
    x = "GPA",
    y = "Count"
  )+
  theme_minimal()
# Will need to model GPA to fix value of 6.0 and NA.
```
## Work Experience Variable
```{r}
summary(admit$work_exp)

admit %>%
  ggplot(aes(x = work_exp)) +
  geom_histogram(fill = "darkblue", alpha = .65, bins = 30) +
  labs(
    title = "Admissions Work Experience",
    x = "Work Experience",
    y = "Count"
  )+
  theme_minimal()
# Clear outlier at 100 Will need to deal with that 
```

## Test Score Variable
```{r}
summary(admit$test_score)


admit %>%
  ggplot(aes(x = test_score)) +
  geom_histogram(fill = "darkblue", alpha = .65, bins = 30) +
  labs(
    title = "Admissions Test Scores",
    x = "Test Score",
    y = "Count"
  )+
  theme_minimal()
# Clear groups of test scores.
# One NA value to deal with 
# Possible visual - admittance by group of test scores. 
```

## Writing Score Variable
```{r}
summary(admit$writing_score)


admit %>%
  ggplot(aes(x = writing_score)) +
  geom_histogram(fill = "darkblue", alpha = .65, bins = 30) +
  labs(
    title = "Admissions Writing Scores",
    x = "Writing Score",
    y = "Count"
  )+
  theme_minimal()
# A couple of outlier scores. Could be input incorrectly, might model to adjust. 
```

## Gender Variable
```{r}
table(admit$gender)

# Two Missing values and a -1 value
# Assume 0 is male and 1 is female.
# -1 is non-binary
# NA means student chose not to reveal gender

admit$gender[admit$gender == 1] = "Female"
admit$gender[admit$gender == 0] = "Male"
admit$gender[admit$gender == -1] = "Non-Binary"
admit$gender[is.na(admit$gender)] = "Declined to Answer"
admit$gender = as.factor(admit$gender)

```

## Volunteer Level Variable
```{r}
summary(admit$volunteer_level)


admit %>%
  ggplot(aes(x = volunteer_level)) +
  geom_histogram(fill = "darkblue", alpha = .65, bins = 30) +
  labs(
    title = "Admissions Volunteer Levels",
    x = "Volunteer Level",
    y = "Count"
  )+
  theme_minimal()
# Viewing histogram it is clear variable should be a factor.  
admit$volunteer_level = as.factor(admit$volunteer_level)
```

## Model GPA to fill missing value
```{r}
admit_model = na.exclude(admit)

set.seed(1212)
k <- 10 

folds <- sample(1:k, nrow(admit_model), replace = TRUE)

# Place holder for errors
cv.errors <- c() 

# write a for loop that performs cross-validation
for (i in 1:k) {
  dat <- admit_model[folds != i, ]
  cv_mod <- glm(gpa ~ ., data = admit_model)
  pred <- as.numeric(predict(cv_mod, admit_model[folds == i, ], id = i, na.rm = T))
  cv.errors[i] <- mean((admit_model$gpa[folds == i] - pred)^2, na.rm = T)
}
# View CV error
mean(cv.errors)

gpa_model = glm(gpa ~ ., data = admit_model)

## Replace values
admit$gpa[is.na(admit$gpa) | admit$gpa > 4] = round(predict(gpa_model, admit[is.na(admit$gpa) | admit$gpa > 4,]),2)


```

## Model work experience to fix outlier
```{r}
admit_model = na.exclude(admit) %>%
  
  # Better for modeling missing value
  filter( work_exp < 100 & decision == "Waitlist") %>%
  select(-decision,-state)

set.seed(123)

folds <- sample(1:k, nrow(admit_model), replace = TRUE)

# Place holder for errors
cv.errors <- c() 

# write a for loop that performs cross-validation
for (i in 1:k) {
  dat <- admit_model[folds != i, ]
  cv_mod <- glm(work_exp ~ . , data = admit_model)
  pred <- as.numeric(predict(cv_mod, admit_model[folds == i, ], id = i, na.rm = T))
  cv.errors[i] <- mean((admit_model$work_exp[folds == i] - pred)^2, na.rm = T)
}
# View CV error
mean(cv.errors)

work_exp_model = glm(work_exp ~ ., data = admit_model)

## Replace values
admit$work_exp[admit$work_exp > 20] = round(predict(work_exp_model, admit[admit$work_exp > 20,]),1)


```

## Model test scores to fix missing value
```{r}
admit_model = na.exclude(admit) %>%
  
  # Better for modeling missing value
  filter(work_exp <= 2 & gpa >=  3.5) %>% 
  select(-decision)


set.seed(1212)
k = 5
folds <- sample(1:k, nrow(admit_model), replace = TRUE)

# Place holder for errors
cv.errors <- c() 

# write a for loop that performs cross-validation
for (i in 1:k) {
  dat <- admit_model[folds != i, ]
  cv_mod <- glm(test_score ~ . , data = admit_model)
  pred <- as.numeric(predict(cv_mod, admit_model[folds == i, ], id = i, na.rm = T))
  cv.errors[i] <- mean((admit_model$test_score[folds == i] - pred)^2, na.rm = T)
}
# View CV error
mean(cv.errors)

test_score_model = glm(test_score ~ ., data = admit_model)

## Replace values
admit$test_score[is.na(admit$test_score)] = round(predict(test_score_model, admit[is.na(admit$test_score),2:8]),0)
```

## Model writing scores to fix likely incorrect values 
```{r}
# Two people have extremely low writing scores which are not in line with the rest of the metrics.
# Model to fix likely incorrect values
admit_model = na.exclude(admit) %>%
  filter(writing_score >= 20)


set.seed(1212)
k = 5
folds <- sample(1:k, nrow(admit_model), replace = TRUE)

# Place holder for errors
cv.errors <- c() 

# write a for loop that performs cross-validation
for (i in 1:k) {
  dat <- admit_model[folds != i, ]
  cv_mod <- glm(writing_score ~ . , data = admit_model)
  pred <- as.numeric(predict(cv_mod, admit_model[folds == i, ], id = i, na.rm = T))
  cv.errors[i] <- mean((admit_model$writing_score[folds == i] - pred)^2, na.rm = T)
}
# View CV error
mean(cv.errors)

writing_score_model = glm(writing_score ~ ., data = admit_model)

## Replace values
admit$writing_score[admit$writing_score < 20] = round(predict(writing_score_model, admit[admit$writing_score < 20,]),0)


```

## Finally fix admission values based on a decision tree
```{r}
admit_model = na.exclude(admit)


set.seed(1212)
k = 5
folds <- sample(1:k, nrow(admit_model), replace = TRUE)

# Place holder for errors
cv.errors <- c() 

# write a for loop that performs cross-validation
for (i in 1:k) {
  dat <- admit_model[folds != i, ]
  dec_mod = rpart(dat$decision ~ .,
              data=dat,
              method = 'class',
              control=rpart.control(minsplit=2, 
                                    minbucket=1))
  pred <- predict(dec_mod, admit_model[folds == i, ],type = "class")
  cv.errors[i] <- sum(admit_model$decision[folds == i] == pred)/length(pred)
}
# View CV error
mean(cv.errors)
#fancyRpartPlot(dec_mod)


decision_model = rpart(decision ~ .,
              data=admit_model,
              method = 'class',
              control=rpart.control(minsplit=2, 
                                    minbucket=1))

## Replace values
admit$decision[is.na(admit$decision)] = predict(decision_model, admit[is.na(admit$decision),], type = "class")


```

# View histograms with cleaned data

## GPA
```{r}
admit %>%
  ggplot(aes(x = gpa)) +
  geom_histogram(fill = "darkblue", alpha = .65, bins = 20) +
  labs(
    title = "Admissions GPAs",
    x = "GPA",
    y = "Count"
  )+
  theme_minimal()

```

## Work Experience 
```{r}

admit %>%
  ggplot(aes(x = work_exp)) +
  geom_histogram(fill = "darkblue", alpha = .65, bins = 30) +
  labs(
    title = "Admissions Work Experience",
    x = "Work Experience",
    y = "Count"
  )+
  theme_minimal()
 
```

## Test Score Variable
```{r}
admit %>%
  ggplot(aes(x = test_score)) +
  geom_histogram(fill = "darkblue", alpha = .65, bins = 30) +
  labs(
    title = "Admissions Test Scores",
    x = "Test Score",
    y = "Count"
  )+
  theme_minimal()

```

## Writing Score Variable
```{r}
admit %>%
  ggplot(aes(x = writing_score)) +
  geom_histogram(fill = "darkblue", alpha = .65, bins = 30) +
  labs(
    title = "Admissions Writing Scores",
    x = "Writing Score",
    y = "Count"
  )+
  theme_minimal()
```

# Write to .csv file
```{r}
write_csv(admit,"cleaned_admissions_data.csv")


```